Phenomenological Robotics

### 2011-12-26 draft.tex
Added EST, EST-Paladyn, Chemero. Compared everyone with Merleau-Ponty, Van Gelder and Freeman.

Van Gelder's idea of a cognitive system being direcly connected to the world can
be seen as an instance of Merleau-Ponty's \emph{intentional arc}, where
Freeman's results support the validity of the \emph{maximal grip} concept.

>>>>

% 2.D. EST vs. Freeman (discrete basin transitions).
% (Source: our BICA10 paper).
% (Source: our Paladyn paper).
% Note: I didn't make the comparison with the discrete basin transitions from
% Freeman here because I believe it would belong together (and better) in the
% end of the section.

Event Segmentation Theory provides a model of how the human brain segments
perception into a sequence of events~\cite{kurby07,zacks07}. This model sustains
that event segmentation is based on the detection of prediction errors in the
sensory stream. In particular, the human brain is permanently making predictions
and comparing them with the actual outcome. Events are detected whenever a
significant disparity between prediction and outcome is encountered.

Nery and Ventura~\cite{nery2011} address the problem of segmenting a robot's
perception into meaningful events, based on Event Segmentation Theory. In this
approach, the prediction of the environment evolution is done using a dynamical
systems approach, where the predictor system has a delayed coupling with the
environment.

% 2.E. Merleau-Ponty (demand) vs. Chemero
% (Source: Thesis journal, pages 100 to 109, 03/03/2011).

Chemero~\cite{chemero2003} criticizes Gibson's original definition of
affordances, deeming it as deceitful. He defends that the perception of the
environment is always connected to the animal, calling the perceived items
\emph{features of situations}. He points out that are not directly related to
``body scale'', but rather to the animal's ``abilities'', arguing that these are
functional properties of the animal that happened to be historically useful.

Chemero defines organisms as ``ability sets'' (and some abilities depend on
others to function, e.g., walking depends on standing). Which abilities can be
used depend on the current situation. He defends that an ``affordance layout''
exists, and that this layout is changed by ``events''.

% Note: The paragraph below tries to put together everything that was described
% above: Van Gelder's coupling is the intentional arc; Freeman's basis of
% attraction implement the maximal grip. Our work connects events and coupling,
% justifying the corroboration of events to Chemero's affordance layout and its
% changes caused by ``events''. Finally, Chemero's concept of affordances itself
% can be seen as an instance of the intentional arc (opportunities presented by
% the environment that ``call us'' to act on it), but on a different timescale.

Van Gelder's idea of a cognitive system being direcly connected to the world can
be seen as an instance of Merleau-Ponty's \emph{intentional arc}, where
Freeman's results support the validity of the \emph{maximal grip} concept.
The work of Nery and Ventura shows the connection between the segmentation of
perception into events and the eventual coupling of an organism and its
environment. Chemero's idea of events changing the layout of affordances is
corroborated by Event Segmentation Theory. The way Chemero presents the
concept of affordances reinforces the idea of \emph{intentional arc}, closing
the loop.
